Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-19 troll/algs_batch.tex
mark content added from the aistats paper

Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
labeled edges from $E(Y)$. We want to build a predictive model for the labels of the remaining
edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
and $q$ of our generative model. They differ in the approximation guarantees they provide, and in
the class of graphs to which they apply. The first algorithm runs on the graph $G$ and estimates
locally the parameters by their empirical means in the training set, which under some density
assumptions are showed to concentrate around their true values. The second algorithm, on the other
hand, exploits the reduced graph $G''$ and computes a maximum likelihood estimation of the
parameters through a global label propagation approach, without making any density assumption.

@@

While we defer the full proof of \autoref{t:active} to the additional material at the end of this
chapter (\autoref{ssec:troll_proof_bayes}), here we give a sketch of the method used, focusing on
parameter $p$. First, by \autoref{l:bins} on negatively associated random variables (stated
\vpageref{l:bins}), we show that with our choice of $Q$, for $\theta=\frac{2Q}{\alpha}$ and for any
$u$ having enough out neighbors (\ie{} $\NNout(u) \ge \theta$), at least $Q$ edges of $\Nout(u)$
are in the training set with high probability.
We use these $Q$ edges to compute $1-\htr(u)$, the empirical probability of drawing a $+1$-labeled
edge from $\Nout(i)$. We show that it is close to the real probability, which according to equation
\eqref{e:pout} is equal to $\frac{1}{2}\,\left(p_u + \qbar_u\right)$. $\qbar_u$ being a sample mean
of i.i.d.~$[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1
\mu\big(p,\cdot\big) dp$, it concentrates around its expectation $\mu_q$.

@@

The approach leading to \autoref{t:active} needs the graph to be sufficiently dense. At first sight,
having $Q = \Omega(\log|V|)$ training edges per nodes does not sound that like such a stringent
requirement. Consider for instance Facebook ---which is neither signed nor directed though. It has
two billion
users~\bottomfootnote{\url{https://investor.fb.com/investor-news/press-release-details/2017/Facebook-Reports-Second-Quarter-2017-Results/}},
each of them having $155\approx 7.2\log|V|$ friends on average~\autocite{facebookFriend16}. However,
the constant hidden in the $\Omega$ notation, and needed to guarantee a good behavior, is much
larger than $7.2$. We will nonetheless see in the experiments that we can still apply the resulting
algorithm (we call it \usrule{}) with satisfying results. Additionally, and in order to address this
density limitation, we now introduce a second method based on label propagation.

>>>>

\begin{newcontent}
Given $G(Y) =(V,E(Y))$, in the batch setting we have at our disposal a training set \trainset{} of
labeled edges from $E(Y)$. We want to build a predictive model for the labels of the remaining
edges. We present two algorithms to do so, which both compute estimates of the all parameters $p$
and $q$ of our generative model. They differ in the approximation guarantees they provide, and in
the class of graphs to which they apply. The first algorithm runs on the graph $G$ and estimates
locally the parameters by their empirical means in the training set, which under some density
assumptions are showed to concentrate around their true values. The second algorithm, on the other
hand, exploits the reduced graph $G''$ and computes a maximum likelihood estimation of the
parameters through a global label propagation approach, without making any density assumption.
\end{newcontent}

@@

\begin{newcontent}
While we defer the full proof of \autoref{t:active} to the additional material at the end of this
chapter (\autoref{ssec:troll_proof_bayes}), here we give a sketch of the method used, focusing on
parameter $p$. First, by \autoref{l:bins} on negatively associated random variables (stated
\vpageref{l:bins}), we show that with our choice of $Q$, for $\theta=\frac{2Q}{\alpha}$ and for any
$u$ having enough out neighbors (\ie{} $\NNout(u) \ge \theta$), at least $Q$ edges of $\Nout(u)$
are in the training set with high probability.
We use these $Q$ edges to compute $1-\htr(u)$, the empirical probability of drawing a $+1$-labeled
edge from $\Nout(i)$. We show that it is close to the real probability, which according to equation
\eqref{e:pout} is equal to $\frac{1}{2}\,\left(p_u + \qbar_u\right)$. $\qbar_u$ being a sample mean
of i.i.d.~$[0,1]$-valued random variables independently drawn from the prior marginal $\int_0^1
\mu\big(p,\cdot\big) dp$, it concentrates around its expectation $\mu_q$.

@@

The approach leading to \autoref{t:active} needs the graph to be sufficiently dense. At first sight,
having $Q = \Omega(\log|V|)$ training edges per nodes does not sound that like such a stringent
requirement. Consider for instance Facebook ---which is neither signed nor directed though. It has
two billion
users~\bottomfootnote{\url{https://investor.fb.com/investor-news/press-release-details/2017/Facebook-Reports-Second-Quarter-2017-Results/}},
each of them having $155\approx 7.2\log|V|$ friends on average~\autocite{facebookFriend16}. However,
the constant hidden in the $\Omega$ notation, and needed to guarantee a good behavior, is much
larger than $7.2$. We will nonetheless see in the experiments that we can still apply the resulting
algorithm (we call it \usrule{}) with satisfying results. Additionally, and in order to address this
density limitation, we now introduce a second method based on label propagation.
\end{newcontent}

### 2017-09-19 troll/experiments.tex
mark content added from the aistats paper

We considered six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
In \adv{}, a trust-based social network for open source developer, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric and its
resistance to attacks is available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
simple, yet powerful, keyword-based technique using a list of positive and negative words. See
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

@@

Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases such twin edges do not disagree,
\ie{} they have the same sign. In practice, we can use that fact to improve our accuracy at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\sgn(\euv) = \sgn(\evu)$. For clarity, when comparing
our methods with existing approaches, we not use that trick. But afterwards, we show in
\autoref{tab:all_mcc_pp} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729

@@

We perform two other experiments. The first one is to assess the effect of predicting reciprocal
edges by their value if available. As expected, the results in \autoref{tab:twin} demonstrate that
it improves MCC when the network has enough twin edges with low disagreements, like \sla{} and
\epi{}. It has no effect when there are few twin edges, like in \aut{}, \wik{} and \kiw{}. And it is
detrimental if when there are many twin edges that do  not agree enough, like in \adv{}, in which
case it is better to rely solely on the learned model.

@@

In the second, we study the effect of our hypothesis that the labeled edges of the training set are
chosen \uar{}. In two of our datasets (\wik{} and \epi{}), edges come the timestamp at which they
were created. A more realistic way of choosing the training set is therefore, for a given training
size $m$, to let $E_0$ be the $m$ oldest edges, and try to predict the remaining, newest one. This
is a common experimental setting in link prediction, where the goal is to infer which pair of nodes
will be connected in the future~\autocite{linkPredSurvey16}. However, as showed in
\input{ts_mcc}
\autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
the existing ones. It is quite surprising that even when the training set is as large as 90\% of all
edges, the MCC are so low compared with those reported \autoref{tab:all_mcc}. The explanation of
why these two sampling strategies produce so different results can be inferred from
\autoref{tab:troll_early_sampling}. Letting $V_{\mathrm{out}}$ be the subset of node of $V$ with at
least one outgoing edge, and defining similarly $V_{\mathrm{in}}$, we see that the node coverage is
naturally larger when sampling at random. Yet when the training set reach 90\% size, with both
sampling strategies we cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
$V_{\mathrm{out}}$. The difference in MCC is thus explained by the last three columns. It shows, for
testing edges \euv{}, the breakdown between three situations, from the least to the most
informative:
\begin{enumerate}[(i),nosep]
  \item no sampled edge outgoing from $u$ nor incoming to $v$;
  \item some sampled edges, but only either outgoing from $u$ or incoming to $v$; and
  \item some sampled edges, both outgoing from $u$ and incoming to $v$;
\end{enumerate}
When the training set is build \uar{}, the vast majority of testing edges falls into the case (iii).
This is no longer true with temporal training set, where situations (ii) and even (i) are more
common, making prediction more difficult. The difference of repartition is particularly strong in
\wik{}, justifying the better performance in \epi{}.

@@

>>>>

We considered six real-world classification datasets. The first four are \dssn{} widely
used as benchmarks for this
task~(\eg{}\autocites{Leskovec2010}{shahriari2014ranking}{wu2016troll}{OnlineCompletion17}):
\begin{newcontent}
In \adv{}, a trust-based social network for open source developer, a user $u$ can certify another
user $v$ with different degrees of trust: \enquote{Observer}, \enquote{Apprentice} (both of which we
consider negative), \enquote{Journeyer} and \enquote{Master} (both of which we consider
positive).\footnote{We download the 7\thup{} of July, 2014 version from
\url{http://www.trustlet.org/datasets/advogato/}.} A full description of this trust metric and its
resistance to attacks is available in the PhD thesis of the website's creator~\autocite[Section
4]{AdvogatoTrustThesis02}).
\end{newcontent}
In \wik{}, there is an edge from user $u$ to user $v$ if $v$ applies for an admin position and $u$
votes for or against that promotion. In \sla{}, a news sharing and commenting website, member $u$
can tag other members $v$ as friends or foes. Finally, in \epi{}, an online shopping website, user
$v$ reviews products and, based on these reviews, another user $u$ can display whether he considers
$v$ to be reliable or not. In addition to these four datasets, we considered two other \ssn{} where
the signs are inferred automatically, rather than given explicitly by the users.  In
\kiw{}~\cite{wikiedits11}, an edge from Wikipedia user $u$ to user $v$ indicates whether they edited
the same article in a constructive manner or not.\footnote{This is the
\href{http://konect.uni-koblenz.de/networks/wikisigned-k2}{KONECT version of the
\enquote{Wikisigned} dataset}, from which we removed self-loops.} Finally, in the
\aut{}~\cite{kumar2016structure} network, an author $u$ cites another author $v$ by either endorsing
or criticizing $v$'s work. The edge sign is derived by classifying the citation sentiment with a
simple, yet powerful, keyword-based technique using a list of positive and negative words. See
\cite{kumar2016structure} for more details.\footnote{We again removed self-loops and merged
multi-edges which are all of the same sign.}

@@

\begin{newcontent}
Finally, there is a low fraction of reciprocal (or \emph{twin}) edges (\ie{} both $\euv \in E$ and $\evu
\in E$), which is a common mechanism of link formation in directed
networks~\autocites{DirectedReciprocity04}{Reciprocity13}). Moreover, in most cases such twin edges do not disagree,
\ie{} they have the same sign. In practice, we can use that fact to improve our accuracy at no
additional computational cost: when predicting $\euv \in E_{\mathrm{test}}$, if the reciprocal edge
\evu{} is part of the training set, we set $\sgn(\euv) = \sgn(\evu)$. For clarity, when comparing
our methods with existing approaches, we not use that trick. But afterwards, we show in
\autoref{tab:twin} when it can be beneficial.
% more references in  http://www.nature.com/articles/srep02729
\end{newcontent}

@@

\begin{newcontent}
We perform two other experiments. The first one is to assess the effect of predicting reciprocal
edges by their value if available. As expected, the results in \autoref{tab:twin} demonstrate that
it improves MCC when the network has enough twin edges with low disagreements, like \sla{} and
\epi{}. It has no effect when there are few twin edges, like in \aut{}, \wik{} and \kiw{}. And it is
detrimental if when there are many twin edges that do  not agree enough, like in \adv{}, in which
case it is better to rely solely on the learned model.

@@

In the second, we study the effect of our hypothesis that the labeled edges of the training set are
chosen \uar{}. In two of our datasets (\wik{} and \epi{}), edges come the timestamp at which they
were created. A more realistic way of choosing the training set is therefore, for a given training
size $m$, to let $E_0$ be the $m$ oldest edges, and try to predict the remaining, newest one. This
is a common experimental setting in link prediction, where the goal is to infer which pair of nodes
will be connected in the future~\autocite{linkPredSurvey16}. However, as showed in
\autoref{tab:troll_ts_mcc}, this makes the problem much more challenging, both for our methods and
the existing ones. It is quite surprising that even when the training set is as large as 90\% of all
edges, the MCC are so low compared with those reported \autoref{tab:all_mcc}.

\input{ts_mcc}

The explanation of
why these two sampling strategies produce so different results can be inferred from
\autoref{tab:troll_early_sampling}. Letting $V_{\mathrm{out}}$ be the subset of node of $V$ with at
least one outgoing edge, and defining similarly $V_{\mathrm{in}}$, we see that the node coverage is
naturally larger when sampling at random. Yet when the training set reach 90\% size, with both
sampling strategies we cover roughly 90\% of the nodes in $V_{\mathrm{out}}$ and in
$V_{\mathrm{out}}$. The difference in MCC is thus explained by the last three columns. It shows, for
testing edges \euv{}, the breakdown between three situations, from the least to the most
informative:
\begin{enumerate}[(i),nosep]
  \item no sampled edge outgoing from $u$ nor incoming to $v$;
  \item some sampled edges, but only either outgoing from $u$ or incoming to $v$; and
  \item some sampled edges, both outgoing from $u$ and incoming to $v$;
\end{enumerate}
When the training set is build \uar{}, the vast majority of testing edges falls into the case (iii).
This is no longer true with temporal training set, where situations (ii) and even (i) are more
common, making prediction more difficult. The difference of repartition is particularly strong in
\wik{}, justifying the better performance in \epi{}.

@@

\end{newcontent}

### 2017-09-19 troll/related.tex
mark content added from the aistats paper

Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} and its weak version~\autocite{davis1967clustering},
that we will describe with more details in \autoref{sub:bias_balance} \vpageref{sub:bias_balance}.
The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. In this section, we provide an
overview of methods tackling the same \esp{} problem as us. Along the way, we give five of them a
name in small capitals, for they are recent and effective. Therefore, we will compare our approaches
with those methods in the experiments of the next section. At the end, we mention some closely
related variants of the original \esp{} problem.

@@

While we presented methods that operate in the batch setting, and will present an online algorithm
as well in \autoref{s:algsonline}, other works have attacked the \esp{} problem from an active
learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active
setting, we are given a budget of edge labels to observe, and are free to select them the way we
want within $E(Y)$. Again, the goal is to make as few mistakes as possible when predicting the sign
of the remaining edges. These two methods build a spanning subgraph of $G$ and query all its edge,
but they differ in its the construction. \Textcite{Cesa-Bianchi2012a} partition $G$ into stars and
connect them in a tree, whereas \textcite{Cesa-Bianchi2012b} cover the graph with cycles, each
containing one test edge and being queried for the other, while respecting the user specified query
budget.  This relies on a different bias than our generative model, and this will be the subject of
our next chapter.  One might also consider an online version of the problem where the topology is
not known in advance but discovered as prediction are made~\autocite{Gentile2013}. This naturally
increases and the difficulty of the problem, as reflected by the computational cost of the solution
proposed, which is quadratic in $|V|$ \emph{at each prediction}.

>>>>

\begin{newcontent}
Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} and its weak version~\autocite{davis1967clustering},
that we will describe with more details in \autoref{sub:bias_balance} \vpageref{sub:bias_balance}.
The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. In this section, we provide an
overview of methods tackling the same \esp{} problem as us. Along the way, we give five of them a
name in small capitals, for they are recent and effective. Therefore, we will compare our approaches
with those methods in the experiments of the next section. At the end, we mention some closely
related variants of the original \esp{} problem.

@@

While we presented methods that operate in the batch setting, and will present an online algorithm
as well in \autoref{s:algonline}, other works have attacked the \esp{} problem from an active
learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active
setting, we are given a budget of edge labels to observe, and are free to select them the way we
want within $E(Y)$. Again, the goal is to make as few mistakes as possible when predicting the sign
of the remaining edges. These two methods build a spanning subgraph of $G$ and query all its edge,
but they differ in its the construction. \Textcite{Cesa-Bianchi2012a} partition $G$ into stars and
connect them in a tree, whereas \textcite{Cesa-Bianchi2012b} cover the graph with cycles, each
containing one test edge and being queried for the other, while respecting the user specified query
budget.  This relies on a different bias than our generative model, and this will be the subject of
our next chapter.  One might also consider an online version of the problem where the topology is
not known in advance but discovered as prediction are made~\autocite{Gentile2013}. This naturally
increases and the difficulty of the problem, as reflected by the computational cost of the solution
proposed, which is quadratic in $|V|$ \emph{at each prediction}.
\end{newcontent}
